Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.Rmd
update shape of non-linearity section

The power comparison illustrates that appropriate conventional test will reject $H_0$ more aggressively than visual tests. In this section, we explore how often they agree with each other by investigating the rejections for the two studies based on $p$-values for each lineup.

@@

For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% 
and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. There are no lineups that are rejected by the visual test but not by the conventional test.

For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% 
and visual tests reject `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the tests rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test.

@@

@@

### Shape of non-linearity

@@

### Shape of heteroskedasticity

>>>>

The power comparison illustrates that appropriate conventional tests will reject $H_0$ more aggressively than visual tests. In this section, we explore how often they agree with each other by investigating the rejections for the two model designs based on $p$-values for each lineup.

@@

For lineups containing non-linearity patterns, conventional tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% and visual tests reject `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% of the time. Of the lineups rejected by the conventional test, `r p_value_cmp_dat %>% filter(type == "non-linearity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by the visual test, that is, approximately half as many as the conventional test. There are no lineups that are rejected by the visual test but not by the conventional test.

In terms of lineups containing heteroskedasticity patterns, `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(conv_reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by conventional tests, while `r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`% are rejected by visual tests. When the conventional test rejects a lineup, there is a great chance (`r p_value_cmp_dat %>% filter(type == "heteroskedasticity") %>% filter(conv_reject == "Reject") %>% count(reject) %>% summarise(pct=round(n[1]/(n[1]+n[2])*100, 0))`%) the visual test will also reject it.

@@

## Shape of non-linearity

Shape of the non-linearity pattern is another factor contributes to the non-linearity model. With different non-linearity patterns, it is expected that the visual test will respond variously.

According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have almost identical power in testing the "U", the "S" and the "M" shape, except for testing the "Triple-U" shape, which have significantly lower power. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969). However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will have little difference to other shapes. The recommendation of the polynomial power for the RESET should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.

For visual tests, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape. Interestingly, the "M" shape is as easy to be detected as the "U" shape, which implies the readability of the shape do not strictly follow the degree of the polynomials.

```{r poly-power-uniform-j, fig.height = 4, fig.cap = "Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours."}

poly_conv_sim %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(RESET4_p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Conventional") %>%
  bind_rows(vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform", type == "polynomial") %>%
  mutate(log_effect_size = log(effect_size)) %>%
  mutate(reject = as.numeric(p_value < 0.05)) %>%
  mutate(shape = c("U", "S", "M", "Triple-U")[shape]) %>%
  mutate(conv_or_not = "Visual")) %>%
  mutate(shape = factor(shape, levels = c("U", "S", "M", "Triple-U"))) %>%
  select(shape, log_effect_size, reject, conv_or_not) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, reject, col = shape), 
              se = FALSE, 
              method = "glm",
              method.args = list(family = binomial())) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  scale_color_manual(values = rcartocolor::carto_pal(4, "TealGrn")) +
  theme_light() +
  theme(legend.position = "bottom") +
  facet_wrap(~conv_or_not) +
  labs(col = "Non-linearity shape")
```

## Shape of heteroskedasticity

@@

### Shape of non-linearity (old)

@@

### Shape of heteroskedasticity (old)

### 2023-01-30 paper_comparison/paper_comparison/paper_comparison.tex
update shape of non-linearity section

The power comparison illustrates that appropriate conventional test will
reject \(H_0\) more aggressively than visual tests. In this section, we
explore how often they agree with each other by investigating the
rejections for the two studies based on \(p\)-values for each lineup.

@@

For lineups containing non-linearity patterns, conventional tests reject
69\% and visual tests reject 32\% of the time. Of the tests rejected by
the conventional test, 46\% are rejected by the visual test, that is,
approximately half as many as the conventional test. There are no
lineups that are rejected by the visual test but not by the conventional
test.

For lineups containing non-linearity patterns, conventional tests reject
76\% and visual tests reject 56\% of the time. Of the tests rejected by
the conventional test, 73\% are rejected by the visual test, that is,
approximately half as many as the conventional test.

@@

@@

\hypertarget{shape-of-non-linearity}{%
\subsubsection{Shape of non-linearity}\label{shape-of-non-linearity}}

@@

\hypertarget{shape-of-heteroskedasticity}{%
\subsubsection{Shape of
heteroskedasticity}\label{shape-of-heteroskedasticity}}

>>>>

The power comparison illustrates that appropriate conventional tests
will reject \(H_0\) more aggressively than visual tests. In this
section, we explore how often they agree with each other by
investigating the rejections for the two model designs based on
\(p\)-values for each lineup.

@@

For lineups containing non-linearity patterns, conventional tests reject
69\% and visual tests reject 32\% of the time. Of the lineups rejected
by the conventional test, 46\% are rejected by the visual test, that is,
approximately half as many as the conventional test. There are no
lineups that are rejected by the visual test but not by the conventional
test.

In terms of lineups containing heteroskedasticity patterns, 76\% are
rejected by conventional tests, while 56\% are rejected by visual tests.
When the conventional test rejects a lineup, there is a great chance
(73\%) the visual test will also reject it.

@@

\hypertarget{shape-of-non-linearity}{%
\subsection{Shape of non-linearity}\label{shape-of-non-linearity}}

Shape of the non-linearity pattern is another factor contributes to the
non-linearity model. With different non-linearity patterns, it is
expected that the visual test will respond variously.

According to Figure \ref{fig:poly-power-uniform-j}, conventional tests
have almost identical power in testing the ``U'', the ``S'' and the
``M'' shape, except for testing the ``Triple-U'' shape, which have
significantly lower power. To understand why this is, one needs to
return to the way the RESET test is applied. It requires a parameter
indicating degree of fitted values to test for, and the recommendation
is to generically use four (\citet{ramsey_tests_1969}). However, the
``Triple-U'' shape constructed from the Hermite polynomials use power up
to 18. If the RESET test had been applied using a higher power no less
than six, the power curve of ``Triple-U'' shape will have little
difference to other shapes. The recommendation of the polynomial power
for the RESET should be revised, perhaps. This illustrates the
sensitivity of conventional testing to the parameters, and it also
points to a limitation that one needs to know the data structure in
order to set the parameters for the test.

For visual tests, we expect the ``U'' shape will be the easiest one to
be detected by subjects followed by the ``S'', ``M'' and ``Triple-U''
shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed
that the power curves are mostly aligned with our expectation, except
for the ``M'' shape. Interestingly, the ``M'' shape is as easy to be
detected as the ``U'' shape, which implies the readability of the shape
do not strictly follow the degree of the polynomials.

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/poly-power-uniform-j-1}

\caption{Power of conventional tests and visual tests on lineups containing four different non-linearity shapes. Power curves with higher order of non-linearity are drawn with deeper colours.}\label{fig:poly-power-uniform-j}
\end{figure}

\hypertarget{shape-of-heteroskedasticity}{%
\subsection{Shape of
heteroskedasticity}\label{shape-of-heteroskedasticity}}

@@

\hypertarget{shape-of-non-linearity-old}{%
\subsubsection{Shape of non-linearity
(old)}\label{shape-of-non-linearity-old}}

@@

\hypertarget{shape-of-heteroskedasticity-old}{%
\subsubsection{Shape of heteroskedasticity
(old)}\label{shape-of-heteroskedasticity-old}}
